### 2014-09-15 PrivacyProtection-2014-09-09.tex
started ThreatTable with mapping to CofI and Vulnerabilities
added combined diagram for it systems
Partial cleanup up formatting and arrangement of sections.

\section{Privacy}

@@

\subsection{Histrocial Approaches to Privacy}

\subsubsection{Aristotle}

@@

\subsubsection{Warren and Brandeis}

@@

\subsubsection{Fried}

@@

\subsection{Legal Aspects of Privacy}

@@

\subsubsection{European Convention on Human Rights - Article 8}

@@

\subsubsection{EU Data Protection Direvtive}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{personal data} is ``any information relating to an identified
  or identifiable natural person {[}\ldots{}{]} (art. 2a)''
  {[}\hyperref[references]{1}{]}
\item
  \textbf{processing} is ``any operation or set of operations which is
  performed upon personal data {[}\ldots{}{]} (art. 2b)''
  {[}\hyperref[references]{1}{]}
\end{itemize}

@@

(\textbf{Note:} this rule is applicable ``whenever the controller uses
equipment situated within the EU''{[}\hyperref[references]{1}{]}. This
will hold for every e-commerce provider \textbf{inside or outside} of
the EU because the customer's computer is \emph{situated within the EU}
anyways.)

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Transparency}: ``The data subject has the right to be informed
  when his personal data is being processed {[}\ldots{}{]} (art. 10 and
  11)'' {[}\hyperref[references]{1}{]} and he has the right how and by
  whom it is processed. Additionally other speficied requirements have
  to be met, i.e.~giving consent.
\item
  \textbf{Legitimate prupose}: ``Personal data can only be processed for
  specified explicit and legitimate pruposes and may not be processes
  futher in a way incompatible with those pruposes (art. 6b)''
  {[}\hyperref[references]{1}{]}.
\item
  \textbf{Proportionality}: ``Personal data may be processed only
  insofar as it is adequate {[}\ldots{}{]} in relation to {[}its{]}
  pruposes {[}\ldots{}{]}; every reasonable step must be taken to ensure
  that data wich are inaccurate or incomplete {[}\ldots{}{]} are erased
  or rectified (art. 6)'' {[}\hyperref[references]{1}{]}
\end{itemize}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Save Harbor
\item
  Passenger Name Record Aggreement
\end{itemize}

@@

\subsubsection{Implementation of the Data Protection Direvtive}

\paragraph{Germany}

@@

\begin{enumerate}
\itemsep1pt\parskip0pt\parsep0pt
\item
  The BDSG has become effective three years too late, thus the EC filed
  a treaty violation proceeding against Germany
  {[}\hyperref[references]{6}{]}.
\item
  The BDSG does not implement \textbf{independet} supervisory
  authorities. The \emph{Bundesdatenschutzbeauftragter} is subordinate
  to the Ministiry of Interior. Although he is not subject to technical
  oversight (\emph{Fachaufsicht}), he is subject to staff supervision by
  the government (\emph{Rechtsaufsicht} durch die Bundesregierung und
  \emph{Dienstaufsicht} durch das Innenministerium) and budget oversight
  (including approval of employees) by the ministry
  {[}\hyperref[references]{7}{]}. Thus the EC filed a treaty violation
  proceeding against Germany, again {[}\hyperref[references]{6}{]}. In
  March 2010 Germany was found guilty of violation of Directive 95/46 by
  the ECJ {[}\hyperref[references]{6}{]}.
\end{enumerate}

(\textbf{Note:} Germany still violates point 2!)

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Roman Catholic Church:}
  \href{http://de.wikipedia.org/wiki/Anordnung_\%C3\%BCber_den_kirchlichen_Datenschutz}{Anordnung
  über den kirchlichen Datenschutz (KDO)}
\item
  \textbf{German Protestant Churches (Synode der Evangelischen Kirche in
  Deutschland):}
  \href{http://de.wikipedia.org/wiki/Datenschutzgesetz_der_Evangelischen_Kirche_in_Deutschland}{Kirchengesetz
  über den Datenschutz der Evangelischen Kirche in Deutschland}
\end{itemize}

Data subjects have the right to\ldots{}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \ldots{} \textbf{disclosure} of
\item
  wether data about them is stored and processed
\item
  which data is stored and processed
\item
  the data sources
\item
  \ldots{} \textbf{correction} of false data
\item
  \ldots{} \textbf{file complaints} to the supervisory authority
\item
  \ldots{} have data \textbf{deleted} or \textbf{blocked}, but
  controllers can prohibit deletion in favour of blocking
\item
  \ldots{} \textbf{Decline} third party access to the data
\end{itemize}

\paragraph{United Kingdom}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \emph{View the data an organisation holds on them. A `subject access
  request' can be obtained for a nominal fee. As of January 2014, the
  maximum fee is £2 for requests to credit reference agencies, £50 for
  health and educational request, and £10 per individual otherwise,}
\item
  \emph{Request that incorrect information be corrected. If the company
  ignores the request, a court can order the data to be corrected or
  destroyed, and in some cases compensation can be awarded.}
\item
  \emph{Require that data is not used in any way that may potentially
  cause damage or distress.}
\item
  *Require that their data is not used for direct marketing.``*
  {[}\hyperref[references]{3}{]}
\end{itemize}

\paragraph{France}

@@

\paragraph{United States of America}

@@

(\textbf{Note:} \emph{Safe harbor} in general is the legal concept to
regulate that a certain conduct will be deemed
{[}\hyperref[references]{4}{]}, but in germany the term is commonly used
as synonym for the agreement between the EU and the US regarding
Directive 95/46.)

\subsection{Privacy Definition and Taxonomy}

\subsubsection{Defining Privacy}

@@

\begin{center}
We follow Fried in defining privacy as the ability to control data about personal information.
\end{center}

@@

\paragraph{The Seven Types of Privacy}

@@

\begin{enumerate}
\item \textbf{Privacy of the Person.}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Body Characteristics
\item
  Weight
\item
  Height
\item
  Other body measures (shoulder width)
\item
  Fingerprints
\item
  DNA Sequence
\item
  Medical Conitions
\item
  Orthopedic conditions (e.g.~limping)
\item
  Having a cold
\item
  \ldots{}
\end{itemize}
\end{enumerate}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  regular visit at church, backery, doctors
\item
  sexual habits
\item
  political activities
\end{itemize}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Briefgeheimnis (in german Legislation)
\item
  Email contents
\item
  Personal direct communication
\item
  Right to free discussion, i.e.~without third parties listening
\end{itemize}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Telefonnummer
\item
  IP Addresse
\item
  Public-administrative Data (Date of Birth, Melderegister)
\item
  Data held by organizations, like Banks or Insurance Companies
\item
  All data that is stored in online services (facebook)
\end{itemize}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Current feeling: Depression, tiredness, stressed, awake
\item
  Thoughts in general
\end{itemize}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  GPS position tracking
\item
  Location of home address
\end{itemize}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Friends
\item
  Joining of orgnaizations (e.g.~political parties)
\end{itemize}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  citizens want to be unidentifiable
\item
  citizens want certain informations about them to be secret
\item
  technical staff needs more or less unrestricted access to data in
  order to maintain the system
\end{itemize}

@@

\paragraph{Implicit Privacy Type Violation (Figure \ref{figure:Implicit Privacy Violation Matrix})}

%\includegraphics{../diagrams/png/implicit-privacy-violation-matrix.png}
\input{./figures/Implicit-Privacy-Violation-Matrix}

>>>>

\chapter{Privacy Protection}

@@

\section{Histrocial Approaches to Privacy}

\subsection{Aristotle}

@@

\subsection{Warren and Brandeis}

@@

\subsection{Fried}

@@

\section{Legal Aspects of Privacy}

@@

\subsection{European Convention on Human Rights - Article 8}

@@

\subsection{EU Data Protection Direvtive}

@@

\begin{itemize}
\item
  \textbf{personal data} is ``any information relating to an identified or identifiable natural person {[}\ldots{}{]} (art. 2a)''
  {[}\hyperref[references]{1}{]}
\item
  \textbf{processing} is ``any operation or set of operations which is performed upon personal data {[}\ldots{}{]} (art. 2b)''
  {[}\hyperref[references]{1}{]}
\end{itemize}

@@

This rule is applicable ``whenever the controller uses equipment situated within the EU''{[}\hyperref[references]{1}{]}.
This will hold for every e-commerce provider \textbf{inside or outside} of the EU because the customer's computer is \emph{situated within the EU} anyways.

@@

\begin{itemize}
\item
\textbf{Transparency}: ``The data subject has the right to be informed when his personal data is being processed {[}\ldots{}{]} (art. 10 and 11)'' {[}\hyperref[references]{1}{]} and he has the right how and by whom it is processed.
Additionally other speficied requirements have to be met, i.e.~giving consent.
\item
\textbf{Legitimate prupose}: ``Personal data can only be processed for specified explicit and legitimate pruposes and may not be processes futher in a way incompatible with those pruposes (art. 6b)''
{[}\hyperref[references]{1}{]}.
\item
\textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate {[}\ldots{}{]} in relation to {[}its{]} pruposes {[}\ldots{}{]}; every reasonable step must be taken to ensure that data wich are inaccurate or incomplete {[}\ldots{}{]} are erased or rectified (art. 6)'' {[}\hyperref[references]{1}{]}
\end{itemize}

@@

\begin{itemize}
\item
  Save Harbor
\item
  Passenger Name Record Aggreement
\end{itemize}

@@

\subsection{Implementation of the Data Protection Direvtive}

\subsubsection*{Germany}

@@

\begin{enumerate}
\item
The BDSG has become effective three years too late, thus the EC filed a treaty violation proceeding against Germany {[}\hyperref[references]{6}{]}.

\item
The BDSG does not implement \textbf{independet} supervisory authorities. 
The \emph{Bundesdatenschutzbeauftragter} is subordinate to the Ministiry of Interior. 
Although he is not subject to technical oversight (\emph{Fachaufsicht}), he is subject to staff supervision by the government (\emph{Rechtsaufsicht} durch die Bundesregierung und  \emph{Dienstaufsicht} durch das Innenministerium) and budget oversight (including approval of employees) by the ministry {[}\hyperref[references]{7}{]}. 
Thus the EC filed a treaty violation proceeding against Germany, again {[}\hyperref[references]{6}{]}. 
In March 2010 Germany was found guilty of violation of Directive 95/46 by the ECJ {[}\hyperref[references]{6}{]}.
\end{enumerate}

Germany still violates point 2!

@@

\begin{itemize}

\item
\textbf{Roman Catholic Church:}
\href{http://de.wikipedia.org/wiki/Anordnung_\%C3\%BCber_den_kirchlichen_Datenschutz}{Anordnung  über den kirchlichen Datenschutz (KDO)}

\item
\textbf{German Protestant Churches (Synode der Evangelischen Kirche in Deutschland):}
\href{http://de.wikipedia.org/wiki/Datenschutzgesetz_der_Evangelischen_Kirche_in_Deutschland}{Kirchengesetz  über den Datenschutz der Evangelischen Kirche in Deutschland}
\end{itemize}

Data subjects have the following rights.

\begin{itemize}
\item
Right to disclosure of wether data about them is stored and processed, which data is stored and processed and the data sources.
\item
Right to correction of false data.

\item
Right to file complaints to the supervisory authority.

\item
Right to have data deleted or blocked, but controllers can prohibit deletion in favour of blocking.

\item
Right to decline third party access to the data.
\end{itemize}

\subsubsection*{United Kingdom}

@@

\begin{itemize}

\item
View the data an organisation holds on them.
A `subject access request' can be obtained for a nominal fee. As of January 2014, the maximum fee is £2 for requests to credit reference agencies, £50 for health and educational request, and £10 per individual otherwise.

\item
Request that incorrect information be corrected.
If the company ignores the request, a court can order the data to be corrected or destroyed, and in some cases compensation can be awarded.

\item
Require that data is not used in any way that may potentially cause damage or distress.

\item
Require that their data is not used for direct marketing. {[}\hyperref[references]{3}{]}
\end{itemize}

\subsubsection*{France}

@@

\subsubsection*{United States of America}

@@

\emph{Safe harbor} in general is the legal concept to regulate that a certain conduct will be deemed {[}\hyperref[references]{4}{]}, but in germany the term is commonly used as synonym for the agreement between the EU and the US regarding Directive 95/46.)

\section{Privacy Definition and Taxonomy}

\subsection{Defining Privacy}

@@

%% TODO %%

* We follow Fried in defining privacy as the ability to control data about personal information.

@@

\subsection{The Seven Types of Privacy}

@@

\begin{enumerate}

\item \textbf{Privacy of the Person.}

@@

Examples include:
\begin{itemize}
\item Body Characteristics: weight, height, sholder width, \ldots
\item Biometric Properties: Fingerprints, DNA Sequence, \ldots
\item Medical Conitions: Orthopedic conditions (e.g.~limping), Having a cold, \ldots
\end{itemize}

\item \textbf{Privacy of Behaviour and Action}

Examples include:
\begin{itemize}
\item regular visit at church, backery, doctors
\item sexual habits
\item political activities
\end{itemize}

\item \textbf{Privacy of Communication}

Examples include:
\begin{itemize}
\item Secrecy of letters ('Briefgeheimnis' in german Legislation).
\item Email contents
\item Personal direct communication
\item Right to free discussion, i.e.~without third parties listening
\end{itemize}

\item \textbf{Privacy of Data and Image}

Examples include:
\begin{itemize}
\item Phone number
\item IP address
\item Public-administrative Data (Date of Birth, Melderegister)
\item Data held by organizations, like Banks or Insurance Companies
\item All data that is stored in online services (facebook)
\end{itemize}

\item \textbf{Privacy of Thoughts and Feelings}

Examples include:
\begin{itemize}
\item Current feeling: Depression, tiredness, stressed, awake
\item Thoughts in general.
\end{itemize}

\item \textbf{Privacy of Location and Space}

Examples include:
\begin{itemize}
\item GPS position tracking
\item Location of home address
\end{itemize}

\item \textbf{Privacy of Association}

Examples include:
\begin{itemize}
\item Friends
\item Joining of orgnaizations (e.g.~political parties)
\end{itemize}
\end{enumerate}

\subsection{Implicit Privacy Type Violation}

Cf. Figure \ref{figure:Implicit Privacy Violation Matrix}

%\includegraphics{../diagrams/png/implicit-privacy-violation-matrix.png}
\input{./figures/Implicit-Privacy-Violation-Matrix}
\input{./figures/ThreatTable}

@@

@@

@@

@@

@@

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt

\item Citizen want to exercise their legal right to privacy. I.e. their right to control information about themselves.
- citizens want to be unidentifiable
- citizens want certain informations about them to be secret
\item Citizen wants to prevent abuse of their private data, and the related negative consequences.
\item
  technical staff needs more or less unrestricted access to data in
  order to maintain the system
\end{itemize}

@@

### 2014-09-15 figures/ThreatTable.tex
Partial cleanup up formatting and arrangement of sections.

\ThreatTableContentRow
{\textbf{(Zero Day) Exploits}
\\A criminal uses (Zero Day) Exploits to obtain access to hardware or software which stores or processes
privacy sensitive data in order to get that data.}
{Criminals want to obtain access or data for personal profit, i.e. by (re-)selling the data or  by processing it themselves. 
However, criminals may have no financial interest, they could also gain personal (ego) profit by testing proof-of-concept attacks.
\\\textbf{Criminal vs Citizen}
\\Citizens only allowed Local Authorities to use their data. They want their data to be secret to others. (Additionally, citizens are
also interested in a working system, which they payed for via taxes.)
\\\textbf{Criminal vs Local Authorities}
\\Local Authorities have capital and reputation invested in a working system. Successful attacks undermine both.
\\\textbf{Criminal vs Maintenance Staff}
\\Staff members have a professional ethos and a duty to provide working systems. Successful attacks offend the former and obstruct the latter.}
{Unsecured Hardware- or Software-Interfaces}
{\textbf{Explicit:} 1,4,6,7}

>>>>

\ThreatTableContentRow
{\textbf{(Zero Day) Exploits}
\\A criminal uses (Zero Day) Exploits to obtain access to hardware or software which stores or processes
privacy sensitive data in order to get that data.}
{Criminals want to obtain access or data for personal profit, i.e. by (re-)selling the data or  by processing it themselves. 
However, criminals may have no financial interest, they could also gain personal (ego) profit by testing proof-of-concept attacks.
\\\textbf{Criminal vs Citizen}
\\Citizens only allowed Local Authorities to use their data. They want their data to be secret to others. (Additionally, citizens are
also interested in a working system, which they payed for via taxes.)
\\\textbf{Criminal vs Local Authorities}
\\Local Authorities have capital and reputation invested in a working system. Successful attacks undermine both.
\\\textbf{Criminal vs Maintenance Staff}
\\Staff members have a professional ethos and a duty to provide working systems. Successful attacks offend the former and obstruct the latter.
\\Staff member employed by L+G or contractor. The Business operation is thretened by loss of confidential data.
}
{Unsecured Hardware- or Software-Interfaces}
{\textbf{Explicit:} 1,4,6,7}
